---
title: "Titanic"
author: "Raian Rith"
date: "2/20/2022"
output: html_document
---

First, we read in the files and use stringsAsFactors = FALSE since we do not want to convert strings to factors for further data cleaning. We then combine both the files so that we can clean them together. 

```{r}
titanic.train <- read.csv("train.csv", stringsAsFactors = FALSE)
titanic.test <- read.csv("test.csv", stringsAsFactors = FALSE)
```

Add a column to the train set to identify that it is from training data set. This will help when we join the two datasets. 

```{r}
titanic.train$IsTrainSet <- TRUE
titanic.test$IsTrainSet <- FALSE
```

To combine we need to make sure all headers are same and that number of columns of train set = number of columns of test set. Headers need to be spelled exactly the same. 

There is no column called Survived in the test set since that is what we have to find. So we create it. 

```{r}
titanic.test$Survived <- NA
```

Now, we join with rbind

```{r}
titanic.full <- rbind(titanic.train, titanic.test)
```

Next, we clean the data in a way so that we can build a model. We deal with the NA's on a column to column basis. 

```{r}
summary(titanic.full)
```


For Embarked, there are 2 missing values. We replace them with the mode of Embarked which is S. 

```{r}
table(titanic.full$Embarked)

titanic.full[titanic.full$Embarked=='', "Embarked"] #Returns the 2 Null values
titanic.full[titanic.full$Embarked=='', "Embarked"] <- 'S' #Replaces the null values with the mode of Embarked (S)

table(titanic.full$Embarked)
```

For Age, there is a lot of missing values. What is the best way of cleaning it ? For now, we replace it with the median. 

```{r}
table(is.na(titanic.full$Age))

age.median <- median(titanic.full$Age, na.rm = TRUE)

titanic.full[is.na(titanic.full$Age), "Age"] <- age.median #replace null with median

table(is.na(titanic.full$Age))
```

For Fare, we have 1 missing value. 

```{r}
table(is.na(titanic.full$Fare))
```


