---
title: "Titanic"
author: "Raian Rith"
date: "2/20/2022"
output: html_document
---

First, we read in the files and use stringsAsFactors = FALSE since we do not want to convert strings to factors for further data cleaning. We then combine both the files so that we can clean them together. 

```{r}
titanic.train <- read.csv("train.csv", stringsAsFactors = FALSE)
titanic.test <- read.csv("test.csv", stringsAsFactors = FALSE)
```

Add a column to the train set to identify that it is from training data set. This will help when we join the two datasets. 

```{r}
titanic.train$IsTrainSet <- TRUE
titanic.test$IsTrainSet <- FALSE
```

To combine we need to make sure all headers are same and that number of columns of train set = number of columns of test set. Headers need to be spelled exactly the same. 

There is no column called Survived in the test set since that is what we have to find. So we create it. 

```{r}
titanic.test$Survived <- NA
```

Now, we join with rbind

```{r}
titanic.full <- rbind(titanic.train, titanic.test)
```

Next, we clean the data in a way so that we can build a model. We deal with the NA's on a column to column basis. 

```{r}
summary(titanic.full)
```


For Embarked, there are 2 missing values. We replace them with the mode of Embarked which is S. 

```{r}
table(titanic.full$Embarked)

titanic.full[titanic.full$Embarked=='', "Embarked"] #Returns the 2 Null values
titanic.full[titanic.full$Embarked=='', "Embarked"] <- 'S' #Replaces the null values with the mode of Embarked (S)

table(titanic.full$Embarked)
```

For Age, there is a lot of missing values. What is the best way of cleaning it ? For now, we replace it with the median. 

```{r}
table(is.na(titanic.full$Age))

age.median <- median(titanic.full$Age, na.rm = TRUE)

titanic.full[is.na(titanic.full$Age), "Age"] <- age.median #replace null with median

table(is.na(titanic.full$Age))
```

For Fare, we have 1 missing value. We will later use regression to find missing fare  but for now we will use the median. 

```{r}
fare.median <- median(titanic.full$Fare, na.rm = TRUE)

titanic.full[is.na(titanic.full$Fare), "Fare"] <- fare.median #replace null with median

```


Next, we need to do some categorical casting. We need to factor Pclass, Sex, and Embarked. Do we factor SibSp (sibling spouse) and Parch (Parent Child)?

```{r}
# str(titanic.full)

titanic.full$Pclass <- as.factor(titanic.full$Pclass)
titanic.full$Sex <- as.factor(titanic.full$Sex)
titanic.full$Embarked<- stas.factor(titanic.full$Embarked)
```



Now that the date has been cleaned, we need to split it back to training and test data sets.

```{r}
titanic.train <- titanic.full[titanic.full$IsTrainSet == TRUE,]
titanic.test <- titanic.full[titanic.full$IsTrainSet == FALSE,]
```

Now we need to do categorical casting on Survived. We could not do that before the split because then we would have 3 factors (including the ones for the test set which we set as NA)

```{r}
titanic.train$Survived<- as.factor(titanic.train$Survived)

```



WE NEED TO DO PARAMETER TUNING, CROSS VALIDATION, 70/30 SPLIT


Now, we need to choose which variables we use as predictors for predicting the binary classification model. 

How to define on R, what to use.  We make a formula in r with the equation we want to use to predict.

```{r}
survived.equation <- "Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked"
survived.formula <- as.formula(survived.equation)
```

We use random forest algorithm for prediction. 

```{r}
#install.packages("randomForest")

library(randomForest)

titanic.model <- randomForest(formula = survived.formula, data = titanic.train, ntree = 500, mtry = 3, nodesize = 0.1 * nrow(titanic.test))
```

Next we specify features. We need to do this bc we want to use passenger id. 

```{r}
features.equation <- "Pclass + Sex + Age + SibSp + Parch + Fare + Embarked"
Survived <- predict(titanic.model, newdata = titanic.test)
```

To submit, we need passengerId and survived. 

```{r}
PassengerId <- titanic.test$PassengerId
output.df <- as.data.frame(PassengerId)
output.df$Survived <- Survived

write.csv(output.df, file = "kaggle_submission.csv", row.names = FALSE)
```

